Oregon 2013 Medication for Asthma Inhalers

### 2017-06-22 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
library(data.table)
library(readxl) 
library(readr)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)

@@

```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
asthma_ndc <- fread(read_path6, showProgress = T)

@@

@@

>>>>

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
library(data.table)
library(readxl) 
library(readr)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
library(broom)

@@

```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
asthma_ndc <- read_csv(read_path6)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

@@

a <- glm(n_obs ~ geo_smk_pm, data = or_ndc_county_time_series, family = "poisson")
summary(a)

```{r}

# create empty dataframe
rr_df <- data.frame(matrix(nrow = 3, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- c("obs_n")

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

meow <- lapply(outcome_list, function(x){

 smk_est <- tidy(glm(as.formula(paste0(x, 
    "~ geo_smk10 + wrf_temp + county + 
    day + offset(log(population))")), 
    or_ndc_county_time_series, family="quasipoisson"))[2, ]

 rr <- c(round(exp(smk_est[1,2]),3),
         round(exp(smk_est[1,2]-1.96*smk_est[1,3]),3),
         round(exp(smk_est[1,2]+1.96*smk_est[1,3]),3))

}) # end apply

for(i in 1:nrow(rr_df)){
  rr_df[i,2:4] <- meow[[i]]
}

# add in a grouping variable
rr_df$group <- NA
rr_df[1:5, 5] <- "Respiratory"
rr_df[6:11, 5] <- "CVD"
rr_df[12, 5] <- "Broken Arm"

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

ggplot(rr_df, aes(x=outcome, y = rel_risk, colour = group)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

@@

### Time series

```{r, eval=FALSE}
# Join data with names of Oregon counties 
oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
# remove the "County" character
# factor(oregon_fips$county)
oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))

```{r time series, warning=FALSE, message=FALSE, echo=FALSE}

oregon_beta_cat_df <- oregon_beta_df %>%
  select(personkey, num_visit, clmid, line, ZIP, dates, index)

# create a working dataset 'or_2013'
oregon_county_beta_cat_ts <- oregon_beta_cat_df %>% 
  # join in the washington county name based on CHARS county identifier
  left_join(or_zip_county, by = "ZIP") %>% 
  # rename 'admit_date_impute' to 'date'
  rename(date = dates) %>%
  # rename(date = admit_date_impute) %>% 
  # group by county of residence and date
  group_by(county, date, index) %>% 
  # sum up each primary diagnosis for each outcome for each day for each county
  summarise(n_obs = n()) %>% 
  # join daily county estimates of pm2.5
  left_join(county_pm, by = c('date', 'county'))

# which(is.na(oregon_county_beta_cat_ts$county))
# oregon_county_beta_cat_ts[113,]
# which(or_zip_county$ZIP=="97078")

# summary(or_ndc_asthma)
# summary(oregon_county_beta_cat_ts)
summary(as.factor(oregon_county_beta_cat_ts$county)) # 125 is missing out of 3592 (3.16%)

# remove NA values
which(is.na(oregon_county_beta_cat_ts$county)) # 7728-7931
oregon_county_beta_cat_ts <- oregon_county_beta_cat_ts[-c(7728:7931),]

write_path5 <- paste0("../data_new/medication/or_county_beta_cat_ts.csv")
write_csv(oregon_county_beta_cat_ts, write_path5)

## Check for missing values in sparsely populated counties. set to 0 (But no)
# check_missing <- filter(oregon_county_beta_cat_ts, is.na(n_obs))
# mutate each to fill in missing values (Still no)
# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
#                                         n_obs:wrf_temp)

plot_ts <- ggplot(oregon_county_beta_cat_ts, aes(y =n_obs, x = date)) +
           facet_wrap(~county, scales = "free") +
           geom_jitter()

```{r}

oregon_beta_rate <- oregon_county_beta_cat_ts %>%
  group_by(county, index) %>%
  summarise(population=n()) %>%  # 143 rows
  mutate

# create empty dataframe
rr_df <- data.frame(matrix(nrow = 12, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- colnames(meta_ts[, which(colnames(meta_ts)=="resp_n"):
    which(colnames(meta_ts)=="broken_arm_n")])

# names 
outcome_names <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
  'Cerebrovascular Disease', 'Heart Failure', 'Ischemic Heart Disease',
  'Myocardial Infarction', 'Broken Arm')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

meow <- lapply(outcome_list, function(x){

 smk_est <- tidy(glm(as.formula(paste0(x, 
    "~ geo_smk10 + wrf_temp + state + county + 
    day + offset(log(population))")), 
    meta_ts, family="quasipoisson"))[2, ]

 rr <- c(round(exp(smk_est[1,2]),3),
         round(exp(smk_est[1,2]-1.96*smk_est[1,3]),3),
         round(exp(smk_est[1,2]+1.96*smk_est[1,3]),3))

}) # end apply

for(i in 1:nrow(rr_df)){
  rr_df[i,2:4] <- meow[[i]]
}

# add in a grouping variable
rr_df$group <- NA
rr_df[1:5, 5] <- "Respiratory"
rr_df[6:11, 5] <- "CVD"
rr_df[12, 5] <- "Broken Arm"

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

ggplot(rr_df, aes(x=outcome, y = rel_risk, colour = group)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
